Predicting Conflict Escalation

### 2023-05-04 docs/hes_final/article.qmd
Reordering Future Work, Ethics, Conclusion, rewrote future research section

While this work did not produce a single best model for conflict escalation predictions, it demonstrated the possibility of such 
predictions from primarily text data. We provided a clear definition of a binary conflict escalation target and described several 
approaches for data aggregation and modeling. We hope this work will contribute to both UNICEF operations and political science 
research. The performance results can serve as a benchmark for future research projects in this 
field. We outline some ideas for future work in the next section.

### Ethical Considerations
This work carries significant ethical considerations, especially if it were to be operationalized. The first and most apparent is bias in the training data. All of our models showed varying performance by region, and overreliance on such models by decisionmakers could lead to resources being diverted away from regions that need them, but which were underrepresented in the training data and thus had less accurate results.

@@

One of the potential research extensions of this work is to continue experimenting with data sampling and aggregation approaches. There 
are many alternatives to the approaches described above, for instance generating daily averages and combining them into $512\times31$ 
(largest number of days in a month) matrices. This could potentially be more sensitive to the signal than the monthly averages and less 
noisy than article level samples.

The transformer models have a great deal of room for hyperparameter tuning. Further experimentation with the size of the encoder output, 
the number of articles per training sample, and implementing overlapping strides during distillation could all improve the downstream 
performance of the classifier. It may also be possible to use the output of the encoder as input to some other classification model 
architecture.

Finally, due to time and resource constraints in this project we used 20% subset of daily news articles. Increasing that percentage could potentially lead to better results.

@@

>>>>

## Ethical Considerations
This work carries significant ethical considerations, especially if it were to be operationalized. The first and most apparent is bias in the training data. All of our models showed varying performance by region, and overreliance on such models by decisionmakers could lead to resources being diverted away from regions that need them, but which were underrepresented in the training data and thus had less accurate results.

@@

Due to time and resource constraints, our work used only 20% of the daily embeddings available through  GDELT, and future research could focus on scaling the methods outline. We also believe that continued experimentation with data sampling and aggregation approaches, as well as model tuning, could yield models that outperform those outlined here without additional data.

Finally, we believe deeper research into regional disparities in model performance is needed. Understanding the limitations of global media coverage for conflict prediction will be critical to any effort to operationalize this work.

@@

While this work did not produce a single best model for conflict escalation predictions, it demonstrated the possibility of such
predictions from primarily text data. We provided a clear definition of a binary conflict escalation target and described several
approaches for data aggregation and modeling. We hope this work will contribute to both UNICEF operations and political science
research. The performance results can serve as a benchmark for future research projects in this
field.
